Multi-Agent Pathfinding with Matching using Increasing Cost Tree Search

### 2021-06-26 docs/paper.tex
Wrote out small numbers

	In practice, joint MDDs are not explicitly constructed but instead, a depth-first search is used to search joint MDDs. This has the benefit of finding a path to the bottom of the joint MDD fast if one exists, while not constructing more of the joint MDD than is necessary for finding a solution, which can get rather large.

@@

	\section{Experimental Results}
	\label{experiments}
	To compare the algorithms and variants described above with each other and with alternative MAPFM algorithms, two types of experiments were performed. In the first, ICTS-based algorithms discussed in this work are compared; in the second, the approximate ordered enumeration variant is compared to other algorithms for MAPFM \cite{donszelmann2021,baauw2021,jong2021,bruin2021}. This comparison data originates in \cite{donszelmann2021}. The values compared are the fraction of instances solved, and the average run time per instance. It should be noticed that in the latter, the recorded run times converge to the 120s timeout used as fewer instances are solved; therefore, while the times are included for completeness at the end of this work (Figure \ref{fig:i-times} and \ref{fig:r-times} for experiments 1 and 2), the fraction of instances solved is guiding in the analysis of results.
	%\newgeometry{left=1cm,right=1cm,top=0.1cm,bottom=0.1cm}
	\begin{figure*}[t]
		\centering
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-1-p}
			\caption{25\% wall, 1 team}
			\label{fig:i-25-1-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-3-p}
			\caption{25\% wall, 3 teams}
			\label{fig:i-25-3-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-1-p}
			\caption{75\% wall, 1 team}
			\label{fig:i-75-1-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-3-p}
			\caption{75\% wall, 3 teams}
			\label{fig:i-75-3-p}
		\end{subfigure}
		\caption{Fraction solved out of 50 maps by different ICTS-based algorithms}
		\label{fig:i-probs}
	\end{figure*}
	\begin{figure*}[b]
		\centering
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-1-p}
			\caption{25\% wall, 1 team}
			\label{fig:r-25-1-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-3-p}
			\caption{25\% wall, 3 teams}
			\label{fig:r-25-3-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-1-p}
			\caption{75\% wall, 1 team}
			\label{fig:r-75-1-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-3-p}
			\caption{75\% wall, 3 teams}
			\label{fig:r-75-3-p}
		\end{subfigure}
		\caption{Fraction solved out of 200 maps by MAPFM algorithms}
		\label{fig:r-probs}
	\end{figure*}

	\restoregeometry
	\subsection{Experimental setup}
	Because MAPFM lies on the fringe of pathfinding research, no standard benchmark exists. For MAPF, the Moving AI benchmarks \cite{sturtevant2012} are often used and potentially a new MAPFM standard benchmark could be defined by generating suitable team-partitions of agents for these MAPF instances; however, it was noticed that optimally solving even small MAPFM instances with comparatively few agents and short distances to goals was already a challenge computationally. This observation informed the creation of a map generator by the author of \cite{jong2021} with various parameters. Using this map generator, two types of 20x20 grids were generated:
	\begin{enumerate}
		\item 25\% wall: a grid with 25\% of tiles being obstacles.
		\item 75\% wall: a grid with 75\% of tiles being obstacles. 
	\end{enumerate}
	For each of these types, maps were generated with between $1$ and $25$ agents either all in the same team or partitioned into $3$ teams as evenly as possible. When solving the instances, for each combination of map-type and number of teams, the number of agents was increased until 0\% of instances could be solved within a timeout of 120s, at which point no more data was recorded for greater numbers of agents.

@@

	\subsection{Experiment 1: ICTS-based MAPFM algorithms}
	To investigate the relative performance of various algorithms discussed in Section \ref{section:icts-matching}, a selection of both exhaustive and ICTS-m variants was made. Their performance in terms of fraction of problems solved within a timeout of 120s is shown in Figure \ref{fig:i-probs}. The meaning of the abbreviations used in the legend are as follows:
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item S: Simple pruning
			\item E: Enhanced pruning
			\item C: Child pruning
		\end{minipage}
		\begin{minipage}{0.45\linewidth}
			\item ID: Independence detection
			\item B: Bounded search
			\item O: Approximate ordered enumeration
		\end{minipage}
	\end{itemize}
	Across scenarios, it is seen that typically exhaustive methods outperform ICTS-m algorithms (\ref{fig:i-probs}), with performance being more similar in the maps with 75\% wall. In particular on 75\% wall maps with 3 teams (\ref{fig:i-75-3-p}), the performance of ICTS-m+ID (and optional pruning) is seen to outperform bounded exhaustive search as $k$ increases, and some of these also manage to solve a small fraction of 75\% maps with 1 team that exhaustive methods cannot within the time-out (\ref{fig:i-75-1-p}).

@@

	The benefit of using ID together with ordered enumeration is less clear. Ordered enumeration with ID performs much better in \ref{fig:i-75-3-p}, a setting with more obstacles and more teams, but ordered enumeration without ID on the other hand is superior in an open setting with a single team (\ref{fig:i-25-1-p}). The overhead associated with the various planning and merging steps (see Algorithm \ref{id-algo}) could be the cause the slowdown observed once $M$ grows large (\ref{fig:i-25-1-p},\ref{fig:i-75-1-p}); once the upper bound is tightened, ICTs only have to be searched shallowly so there is no point in going through the steps of ID. This explains why no speedup is seen, as between ICTS-m and ICTS-m+ID. Another factor is that if all agents belong to the same team, it is somewhat unlikely that there will a significant benefit to using ID, as agents from the same team planned independently often route to the same goal.

	\subsection{Experiment 2: Algorithms for MAPFM}
	From the variants discussed and compared above, the Exh+E+B+O variant was selected for comparison with various other algorithms for MAPFM. This results in the following set of algorithms whose performance is shown in Figure \ref{fig:r-probs}.
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item ICTS with approximate sorted exhaustive matching (Exh+E+B+O)
			\item A*+ID+OD with matching ID and sorted exhaustive matching \cite{bruin2021}
			\item EPEA* with matching ID and sorted exhaustive matching \cite{jong2021}
		\end{minipage}
		\hspace{1cm}
		\begin{minipage}{0.40\linewidth}
			\item M* with sorted exhaustive matching \cite{donszelmann2021}
			\item CBM adapted to use SIC, using the successive shortest path algorithm (SSP) \cite{goldberg1987} as subroutine for solving teams \cite{baauw2021}.
		\end{minipage}
	\end{itemize}
	With exception of CBM, all methods described use a variant of sorted exhaustive matching. This is because from each family of algorithms one of the best performing algorithms was chosen, which always was such an exhaustive method as opposed to a generalization of the base algorithm to MAPFM (ICTS-m in this work). A*+ID+OD and EPEA* also make use of matching ID, which is ID applied to the MAPFM instance using the respective exhaustive algorithm as MAPFM solver. The initial groups are here the teams, which is why in single team scenarios (Subfigures \ref{fig:r-25-1-p},\ref{fig:r-75-1-p}) matching ID has no effect. Hypothetically, matching ID could explain why both A*+ID+OD and EPEA* outperform the methods using just sorted exhaustive matching in maps with 75\% wall and 3 teams (\ref{fig:r-75-3-p}).

@@

	\printbibliography
	\newpage
	\begin{figure*}[t]
		\centering
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-1}
			\caption{25\% wall, 1 team}
			\label{fig:i-25-1}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-3}
			\caption{25\% wall, 3 teams}
			\label{fig:i-25-3}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-1}
			\caption{75\% wall, 1 team}
			\label{fig:i-75-1}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-3}
			\caption{75\% wall, 3 teams}
			\label{fig:i-75-3}
		\end{subfigure}
		\caption{Average solving time of different ICTS-based algorithms with a timeout of 120s}
		\label{fig:i-times}
	\end{figure*}
	\begin{figure*}[b]
		\centering
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-1}
			\caption{25\% wall, 1 team}
			\label{fig:r-25-1}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-3}
			\caption{25\% wall, 3 teams}
			\label{fig:r-25-3}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-1}
			\caption{75\% wall, 1 team}
			\label{fig:r-75-1}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-3}
			\caption{75\% wall, 3 teams}
			\label{fig:r-75-3}
		\end{subfigure}
		\caption{Average solving time of MAPFM algorithms with a timeout of 120s}
		\label{fig:r-times}
	\end{figure*}

>>>>

	In practice, joint MDDs are not explicitly constructed but instead, a depth-first search is used to search joint MDDs. This has the benefit of finding a path to the bottom of the joint MDD fast if one exists, while not constructing more of the joint MDD than is necessary for finding a solution, which can get large.

@@

	\section{Experimental Results}
	\label{experiments}
	To compare the algorithms and variants described above with each other and with alternative MAPFM algorithms, two types of experiments were performed. In the first, ICTS-based algorithms discussed in this work are compared; in the second, the approximate ordered enumeration variant is compared to other algorithms for MAPFM \cite{donszelmann2021,baauw2021,jong2021,bruin2021}. This comparison data originates in \cite{donszelmann2021}. The values compared are the fraction of instances solved, and the average run time per instance. It should be noticed that in the latter, the recorded run times converge to the 120s timeout used as fewer instances are solved; therefore, while the times are included for completeness at the end of this work (Figure \ref{fig:i-times} and \ref{fig:r-times} for experiments 1 and 2), the fraction of instances solved is guiding in the analysis of results.
	%\newgeometry{left=1cm,right=1cm,top=0.1cm,bottom=0.1cm}
	\begin{figure*}[t]
		\centering
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-1-p}
			\caption{25\% wall, one team}
			\label{fig:i-25-1-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-3-p}
			\caption{25\% wall, three teams}
			\label{fig:i-25-3-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-1-p}
			\caption{75\% wall, one team}
			\label{fig:i-75-1-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-3-p}
			\caption{75\% wall, three teams}
			\label{fig:i-75-3-p}
		\end{subfigure}
		\caption{Fraction solved out of 50 maps by different ICTS-based algorithms}
		\label{fig:i-probs}
	\end{figure*}
	\begin{figure*}[b]
		\centering
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-1-p}
			\caption{25\% wall, one team}
			\label{fig:r-25-1-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-3-p}
			\caption{25\% wall, three teams}
			\label{fig:r-25-3-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-1-p}
			\caption{75\% wall, one team}
			\label{fig:r-75-1-p}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-3-p}
			\caption{75\% wall, three teams}
			\label{fig:r-75-3-p}
		\end{subfigure}
		\caption{Fraction solved out of 200 maps by MAPFM algorithms}
		\label{fig:r-probs}
	\end{figure*}

	\restoregeometry
	\subsection{Experimental setup}
	Because MAPFM lies on the fringe of pathfinding research, no standard benchmark exists. For MAPF, the Moving AI benchmarks \cite{sturtevant2012} are often used and potentially a new MAPFM standard benchmark could be defined by generating suitable team-partitions of agents for these MAPF instances; however, it was noticed that optimally solving even small MAPFM instances with comparatively few agents and short distances to goals was already a challenge computationally. This observation informed the creation of a map generator by the author of \cite{jong2021} with various parameters. Using this map generator, two types of 20x20 grids were generated:
	\begin{enumerate}
		\item 25\% wall: a grid with 25\% of tiles being obstacles.
		\item 75\% wall: a grid with 75\% of tiles being obstacles. 
	\end{enumerate}
	For each of these types, maps were generated with between $1$ and $25$ agents either all in the same team or partitioned into three teams as evenly as possible. When solving the instances, for each combination of map-type and number of teams, the number of agents was increased until 0\% of instances could be solved within a timeout of 120s, at which point no more data was recorded for greater numbers of agents.

@@

	\subsection{Experiment 1: ICTS-based MAPFM algorithms}
	To investigate the relative performance of various algorithms discussed in Section \ref{section:icts-matching}, a selection of both exhaustive and ICTS-m variants was made. Their performance in terms of fraction of problems solved within a timeout of 120s is shown in Figure \ref{fig:i-probs}. The meaning of the abbreviations used in the legend are as follows:
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item S: Simple pruning
			\item E: Enhanced pruning
			\item C: Child pruning
		\end{minipage}
		\begin{minipage}{0.45\linewidth}
			\item ID: Independence detection
			\item B: Bounded search
			\item O: Approximate ordered enumeration
		\end{minipage}
	\end{itemize}
	Across scenarios, it is seen that typically exhaustive methods outperform ICTS-m algorithms (\ref{fig:i-probs}), with performance being more similar in the maps with 75\% wall. In particular on 75\% wall maps with three teams (\ref{fig:i-75-3-p}), the performance of ICTS-m+ID (and optional pruning) is seen to outperform bounded exhaustive search as $k$ increases, and some of these also manage to solve a small fraction of 75\% maps with one team that exhaustive methods cannot within the time-out (\ref{fig:i-75-1-p}).

@@

	The benefit of using ID together with ordered enumeration is less clear. Ordered enumeration with ID performs much better in \ref{fig:i-75-3-p}, a setting with more obstacles and more teams, but ordered enumeration without ID on the other hand is superior in an open setting with a single team (\ref{fig:i-25-1-p}). The overhead associated with the various planning and merging steps (see Algorithm \ref{id-algo}) could be the cause the slowdown observed once $M$ grows large (\ref{fig:i-25-1-p},\ref{fig:i-75-1-p}); once the upper bound is tightened, ICTs only have to be searched shallowly so there is no point in going through the steps of ID. This explains why no speedup is seen, as between ICTS-m and ICTS-m+ID. Another factor is that if all agents belong to the same team, it is unlikely that there will a significant benefit to using ID, as agents from the same team planned independently often route to the same goal.

	\subsection{Experiment 2: Algorithms for MAPFM}
	From the variants discussed and compared above, the Exh+E+B+O variant was selected for comparison with various other algorithms for MAPFM. This results in the following set of algorithms whose performance is shown in Figure \ref{fig:r-probs}.
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item ICTS with approximate sorted exhaustive matching (Exh+E+B+O)
			\item A*+ID+OD with matching ID and sorted exhaustive matching \cite{bruin2021}
			\item EPEA* with matching ID and sorted exhaustive matching \cite{jong2021}
		\end{minipage}
		\hspace{1cm}
		\begin{minipage}{0.40\linewidth}
			\item M* with sorted exhaustive matching \cite{donszelmann2021}
			\item CBM adapted to use SIC, using the successive shortest path algorithm (SSP) \cite{goldberg1987} as subroutine for solving teams \cite{baauw2021}.
		\end{minipage}
	\end{itemize}
	With exception of CBM, all methods described use a variant of sorted exhaustive matching. This is because from each family of algorithms one of the best performing algorithms was chosen, which always was such an exhaustive method as opposed to a generalization of the base algorithm to MAPFM (ICTS-m in this work). A*+ID+OD and EPEA* also make use of matching ID, which is ID applied to the MAPFM instance using the respective exhaustive algorithm as MAPFM solver. The initial groups are here the teams, which is why in single team scenarios (Subfigures \ref{fig:r-25-1-p},\ref{fig:r-75-1-p}) matching ID has no effect. Hypothetically, matching ID could explain why both A*+ID+OD and EPEA* outperform the methods using just sorted exhaustive matching in maps with 75\% wall and three teams (\ref{fig:r-75-3-p}).

@@

	\printbibliography
	\newpage
	\begin{figure*}[t]
		\centering
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-1}
			\caption{25\% wall, one team}
			\label{fig:i-25-1}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-3}
			\caption{25\% wall, three teams}
			\label{fig:i-25-3}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-1}
			\caption{75\% wall, one team}
			\label{fig:i-75-1}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-3}
			\caption{75\% wall, three teams}
			\label{fig:i-75-3}
		\end{subfigure}
		\caption{Average solving time of different ICTS-based algorithms with a timeout of 120s}
		\label{fig:i-times}
	\end{figure*}
	\begin{figure*}[b]
		\centering
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-1}
			\caption{25\% wall, one team}
			\label{fig:r-25-1}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-3}
			\caption{25\% wall, three teams}
			\label{fig:r-25-3}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-1}
			\caption{75\% wall, one team}
			\label{fig:r-75-1}
		\end{subfigure}
		\begin{subfigure}{\graphwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-3}
			\caption{75\% wall, three teams}
			\label{fig:r-75-3}
		\end{subfigure}
		\caption{Average solving time of MAPFM algorithms with a timeout of 120s}
		\label{fig:r-times}
	\end{figure*}
